Packaging research artefacts with RO-Crate

### 2022-01-07 response-to-review.md

# Response to peer review, Data Science submission 706-1686

_Response to  open peer review at <https://datasciencehub.net/paper/packaging-research-artefacts-ro-crate> of “Packaging research artefacts with RO-Crate” submitted to Data Science._

### 2022-01-08 latex/ro-crate.tex
DOI as published in Data Science
Alphabetical bibliography as in proof
latex content+bibliography from ro-crate-proof.tex
but with our images

%BeginFileInfo
%%Publisher=IOS
%%Project=DS
%%Manuscript=DS210053
%%MS position=
%%Stage=201
%%TID=imikolaityte
%%Pages=0
%%Format=2019
%%Distribution=vtex
%%Destination=DVI
%%PDF type=
%%DVI.Maker=luadvi
%%PS.Maker=luaps
%%History1=2021.12.28 09:11
%EndFileInfo
%
% Journal: Data Science (DS), IOS Press
% Typeset in LaTeX2e by VTeX Ltd., Vilnius, Lithuania
%
%Spelling_date
%\RequirePackage{float} % when algorithm.sty used in manuscript
\documentclass[ds,v1.1.2,openaccess]{iosart2x}%% ,seceqn
\usepackage{mathbh}
%
%\usepackage{}
\usepackage{tmultirow}
\usepackage{listings}
%
\PROOF
%\PROOF
%\CRC

\volume{-1}
\issue{-1}
\pubyear{2021}
\firstpage{1}
\lastpage{42}
\articletype{research-article} %%
%http://www.vtex.vmt/docs/wiki/index.php/IOS_Press_instrukcijos_%28Kaunas%29#Straipsnio_tipas_.28.5Carticletype.7B.7D.29
\aid{DS210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
\doi{10.3233/DS-210053}% Updated by VTEXPTS2LaTeX.exe, 28.12.2021 09:11
\vtexed{Ingrida}

\startlocaldefs
\setpkgattr{license}{type}{CC@BY} %% CC@BY@NC
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% indentation after theorem-like environments - auto
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\newtheorem*{}{} %% unnumbered
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\theoremstyle{plain}
%%\newtheorem{}{}%[]
%%\theoremstyle{definition}
%%\newtheorem{}{}%[]
\endlocaldefs

@@

@@

\pretitle[XML]{Research Article}
%\pretitle{}
\title{Packaging research artefacts with RO-Crate}
%\title{\thanks{}}
\runtitle{Packaging research artefacts with RO-Crate}
%\subtitle{}
%%% initials without spaces
% {\inits{}\fnms{} \snm{} \suffix{}\ead[label=e?]{}
%\ead[label=p1,phone]{}\ead[label=f1,fax]{}\ead[label=o?,orcid]{}
%\thanks{}}
\begin{aug}
\author[A,B]{\inits{S.}\fnms{Stian} \snm{Soiland-Reyes}\ead
[label=e1]{soiland-reyes@manchester.ac.uk}\ead
[label=o1,orcid]{0000-0001-9842-9718}\thanks{Corresponding author.
\printead{e1}.}}
\author[C]{\inits{P.}\fnms{Peter} \snm{Sefton}\ead
[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}}
\author[D]{\inits{M.}\fnms{Merc\`{e}} \snm{Crosas}\ead
[label=o3,orcid]{0000-0003-1304-1939}}
\author[E]{\inits{L.J.}\fnms{Leyla Jael} \snm{Castro}\ead
[label=o4,orcid]{0000-0003-3986-0510}}
\author[F]{\inits{F.}\fnms{Frederik} \snm{Coppens}\ead
[label=o5,orcid]{0000-0001-6565-5145}}
\author[G]{\inits{J.M.}\fnms{Jos\'{e} M.} \snm{Fern\'{a}ndez}\ead
[label=o6,orcid]{0000-0002-4806-5140}}
\author[H]{\inits{D.}\fnms{Daniel} \snm{Garijo}\ead
[label=o7,orcid]{0000-0003-0454-7145}}
\author[I]{\inits{B.}\fnms{Bj\"{o}rn} \snm{Gr\"{u}ning}\ead
[label=o8,orcid]{0000-0002-3079-6586}}
\author[J]{\inits{M.}\fnms{Marco} \snm{La Rosa}\ead
[label=o9,orcid]{0000-0001-5383-6993}}
\author[K]{\inits{S.}\fnms{Simone} \snm{Leo}\ead
[label=o10,orcid]{0000-0001-8271-5429}}
\author[L]{\inits{E.}\fnms{Eoghan} \snm{\'{O} Carrag\'{a}in}\ead
[label=o11,orcid]{0000-0001-8131-2150}}
\author[M]{\inits{M.}\fnms{Marc} \snm{Portier}\ead
[label=o12,orcid]{0000-0002-9648-6484}}
\author[N]{\inits{A.}\fnms{Ana} \snm{Trisovic}\ead
[label=o13,orcid]{0000-0003-1991-0533}}
\author[O]{\snm{RO-Crate Community}\ead
[label=u1,url]{https://www.researchobject.org/ro-crate/community}}
\author[P]{\inits{P.}\fnms{Paul} \snm{Groth}\ead
[label=e15]{p.t.groth@uva.nl}\ead[label=o15,orcid]{0000-0003-0183-6910}}
\author[Q]{\inits{C.}\fnms{Carole} \snm{Goble}\ead
[label=e16]{carole.goble@manchester.ac.uk}\ead
[label=o16,orcid]{0000-0003-1219-2137}}
%%%%%% printead order: TEL FAX E-MAIL URL
%\thanks{Corresponding author: name surname, address. \printead{}.}
%\thanks{Corresponding author. \printead{}.}
%\thanks{}%% Move support info to acks
\runauthor{S. Soiland-Reyes et al.} %% auto
% {\orgname{} \cny{}\printead[presep={\\}]{}}
\address[A]{Department of Computer Science, \orgname{The University of
Manchester}, \cny{UK}\printead[presep={\\}]{e1,o1}}
\address[B]{Informatics Institute, \orgname{University of Amsterdam},
\cny{The Netherlands}}
\address[C]{Faculty of Science, \orgname{University Technology Sydney},
\cny{Australia}\printead[presep={\\}]{e2,o2}}
\address[D]{Institute for Quantitative Social Science, \orgname{Harvard
University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{o3}}
\address[E]{\orgname{ZB MED Information Centre for Life Sciences},
Cologne, \cny{Germany}\printead[presep={\\}]{o4}}
\address[F]{\orgname{VIB-UGent Center for Plant Systems Biology}, Gent,
\cny{Belgium}\printead[presep={\\}]{o5}}
\address[G]{\orgname{Barcelona Supercomputing Center}, Barcelona, \cny
{Spain}\printead[presep={\\}]{o6}}
\address[H]{Ontology Engineering Group, \orgname{Universidad Polit\'
{e}cnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{o7}}
\address[I]{Bioinformatics Group, Department of Computer Science,
\orgname{Albert-Ludwigs-University Freiburg}, Freiburg, \cny
{Germany}\printead[presep={\\}]{o8}}
\address[J]{\orgname{PARADISEC}, Melbourne, \cny{Australia}\printead
[presep={\\}]{o9}}
\address[K]{\orgname{Center for Advanced Studies, Research, and
Development in Sardinia} (CRS4), Pula (CA), \cny{Italy}\printead
[presep={\\}]{o10}}
\address[L]{\orgname{University College Cork}, \cny{Ireland}\printead
[presep={\\}]{o11}}
\address[M]{\orgname{Vlaams Instituut voor de Zee}, Oostende, \cny
{Belgium}\printead[presep={\\}]{o12}}
\address[N]{Institute for Quantitative Social Science, \orgname{Harvard
University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{o13}}
\address[O]{\printead{u1}~(see Appendix B)}
\address[P]{Informatics Institute, \orgname{University of Amsterdam},
\cny{The Netherlands}\printead[presep={\\}]{e15,o15}}
\address[Q]{Department of Computer Science, \orgname{The University of
Manchester}, \cny{UK}\printead[presep={\\}]{e16,o16}}
\end{aug}
%%%%%%
\begin{review}{editor}
\reviewer{\fnms{Silvio} \snm{Peroni}\ead
[label=oe1,orcid]{0000-0003-0530-4305}\printead[presep={\ }]{oe1}}
%\reviewer{\fnms{} \snm{}\ead[label=oe2,orcid]{}\printead[presep={\
%}]{oe2}}
\end{review}

\begin{review}{solicited}
\reviewer{\fnms{Angelo A.} \snm{Salatino}\ead
[label=os1,orcid]{0000-0002-4763-3943}\printead[presep={\ }]{os1}}
\reviewer{\fnms{Alessia} \snm{Bardi}\ead
[label=os2,orcid]{0000-0002-1112-1292}\printead[presep={\ }]{os2}}
\reviewer{\snm{one anonymous reviewer}}
\end{review}
%
%\dedicated{}
%
% HISTORY:
\received{\sday{13} \smonth{8} \syear{2021}}
\accepted{\sday{23} \smonth{9} \syear{2021}}
%
\begin{abstract}
%\parttitle{}
An increasing number of researchers support reproducibility by
including pointers to and
descriptions of datasets, software and methods in their publications.
However, scientific
articles may be ambiguous, incomplete and difficult to process by
automated systems. In
this paper we introduce RO-Crate, an open, community-driven, and
lightweight approach
to packaging research artefacts along with their metadata in a machine
readable manner.
RO-Crate is based on Schema.org annotations in JSON-LD, aiming to
establish best
practices to formally describe metadata in an accessible and practical
way for their
use in a wide variety of situations.

An RO-Crate is a structured archive of all the items that contributed
to a research
outcome, including their identifiers, provenance, relations and
annotations. As a general
purpose packaging approach for data and their metadata, RO-Crate is
used across multiple
areas, including bioinformatics, digital humanities and regulatory
sciences. By applying
``just enough'' Linked Data standards, RO-Crate simplifies the process of
making research outputs FAIR while also enhancing research reproducibility.

An RO-Crate for this article\footnote{\url
{https://w3id.org/ro/doi/10.5281/zenodo.5146227}}
is archived at \url{https://doi.org/10.5281/zenodo.5146227}.
\end{abstract}
%
\begin{keyword} %% 1 u.c.
\kwd{Data publishing}
\kwd{data packaging}
\kwd{FAIR}
\kwd{Linked Data}
\kwd{metadata}
\kwd{reproducibility}
\kwd{research object}
\end{keyword}
%
\end{frontmatter}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Main text entry area:
\section{Introduction}

@@

\begin{figure}%[t!]
\includegraphics{210053f01}
\caption{Conceptual overview of RO-Crate. A \emph{Persistent
Identifier} (PID) \cite{doi:10.1371/journal.pbio.2001414} points to a
\emph{Research Object} (RO), which may be archived using different
packaging approaches like BagIt \cite{doi:10.17487/rfc8493}, OCFL \cite
{ocfl_2020}, git or ZIP. The RO is described within a \emph{RO-Crate
Metadata File}, providing identifiers for \emph{authors} using ORCID,
\emph{organisations} using Research Organization Registry (ROR) \cite
{doi:10.6087/kcse.192} and licences such as Creative Commons using SPDX
identifiers. The \emph{RO-Crate content} is further described with
additional metadata following a Linked Data approach. Data can be
embedded files and directories, as well as links to external Web
resources, PIDs and nested RO-Crates.}
\label{fig:conceptual}
\end{figure}

@@

\begin{figure}%[t!]
\includegraphics{210053f02}
\caption{Simplified UML class diagram of RO-Crate. The \emph
{RO-Crate Metadata File} conforms to a version of the specification;
and contains a JSON-LD graph \cite{sporny_2014} that describes the
entities that make up the RO-Crate. The \emph{RO-Crate Root Data
Entity} represent the Research Object as a dataset. The RO-Crate
aggregates \emph{data entities} (\texttt{hasPart}) which are further
described using \emph{contextual entities} (which may include
aggregated and non-aggregated data entities). Multiple types and
relations from Schema.org allow annotations to be more specific,
including figures, nested datasets, computational workflows, people,
organisations, instruments and places. Contextual entities not
otherwise cross-referenced from other entities' properties (\emph
{describes}) can be grouped under the root entity (\texttt{mentions}).}
\label{fig:uml}
\end{figure}

@@

%
\begin{figure}%[t]
\includegraphics{210053f03}
\caption{Separation of Concerns in BCO RO-Crate. BioCompute
Object (IEEE2791) is a JSON file that structurally explains the purpose
and implementation of a computational workflow, for instance
implemented in Common Workflow Language (CWL), that installs the
workflow's software dependencies as Docker containers or BioConda
packages. An example execution of the workflow shows the different
kinds of result outputs, which may be external, using GitHub LFS \cite
{github-lfs} to support larger data. RO-Crate gathers all these local
and external resources, relating them and giving individual
descriptions, for instance permanent DOI identifiers for reused
datasets accessed from Zenodo, but also adding external identifiers to
attribute authors using ORCID or to identify which licences apply to
individual resources. The RO-Crate and its local files are captured in
a BagIt whose checksum ensures completeness, combined with Big Data Bag
\cite{doi:10.1109/BigData.2016.7840618} features to ``complete'' the
bag with large external files such as the workflow outputs.}
\label{fig:sep_concerns}
\end{figure}
%
\subsection{Digital humanities: Cultural heritage}% {#culturalheritage}

@@

\begin{figure}%[t!]
\includegraphics{210053f04}
\caption{One aspect of Harvard Data Commons. Automatic
encapsulation and deposit of artefacts from data management tools used
during active research at the Harvard Dataverse repository.}
\label{fig:hdc}
\end{figure}

@@

% structpyb loaded by imikolaityte, 2021-12-28 10:21:28
%
\begin{thebibliography}{}

@@

\end{thebibliography}
\end{document}
%

>>>>

%% ro-crate paper
%% (c) 2020-2021 The University of Manchester
%% 
%% SPDX-License-Identifier: CC-BY-4.0

% add. options: [seceqn,secthm,crcready]
\documentclass[ds,crcready]{iosart2x}

%%%%%%%%%%% Put your definitions here

% From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
\usepackage{fontspec}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{xcolor}
\definecolor{darkerblue}{cmyk}{1,1,0,0.45}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    citecolor=black,
    urlcolor=darkerblue}

\usepackage{unicode-math}
\usepackage{graphicx}

% Hyperlinks on DOI permalinks – just 20 years after DOI invention!
% 
% We had to make our own copy of 
% https://ctan.org/pkg/doi?lang=en 
% where commands are renamed to \doix
% because iosart2x.cls and ios1.bst have 
% a different use of \doiurl \doi etc.
% We also set the doixtext prefix empty
% to avoid doi:doi:
\usepackage{doix}
\renewcommand{\doixtext}{}
\renewcommand{\doiurl}[1]{\doix{#1}}

\DeclareGraphicsExtensions{.pdf,.png,.svg}
\graphicspath{../content/images/}
\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
\def\markdownOptionOutputDir{../output}
\def\markdownOptionHeaderAttributes{true}

% for table resizing
\usepackage{booktabs}
\usepackage{multirow}

%%%%%%%%%%% End of definitions

\pubyear{2022}
\volume{0}
\firstpage{1}
\lastpage{1}
\def\doi{10.3233/DS-210053}

@@

% Welcome to the 2000s with UNICODE fonts!
% Adapted from
% https://www.overleaf.com/learn/latex/Articles%2FOpenType-based_math_typesetting:_An_introduction_to_the_STIX2_OpenType_fonts
\setmathfont{STIX2Math}[
    Extension={.otf},
    Path=./STIX2fonts/,
    Scale=1
]
\setmainfont{STIX2Text}[
    Extension={.otf},
    Path=./STIX2fonts/,
    UprightFont={*-Regular},
    BoldFont={*-Bold},
    ItalicFont={*-Italic},
    BoldItalicFont={*-BoldItalic}
]

% https://typeof.net/Iosevka/ mono font has 
% all the logic symbols we need
%\setmonofont{iosevka-slab}[
%    Extension={.ttf},
%    Path=./ttf-iosevka-slab-6.1.3/,
%    Scale=MatchLowercase,
%    UprightFont={*-light},
%    ItalicFont={*-lightitalic}
%]

@@

%\pretitle{}
\title{Packaging research artefacts with RO-Crate}
\runtitle{Packaging research artefacts with RO-Crate}
%\subtitle{}

% For one author:
%\author{\inits{N.}\fnms{Name1} \snm{Surname1}\ead[label=e1]{first@somewhere.com}\ead[label=o1,orcid]{0000-0000-0000-0001}}
%\address{Department first, \orgname{University or Company name},
%Abbreviate US states, \cny{Country}\printead[presep={\\}]{e1,o1}}

% Two or more authors:
\author[SSR1,SSR2]{\inits{S.}\fnms{Stian}
    \snm{Soiland-Reyes}\ead[label=eSSR]{soiland-reyes@manchester.ac.uk}\ead[label=oSSR,orcid]{0000-0001-9842-9718}%
    \thanks{Corresponding author. \printead{eSSR}.}}, 
\author[PS]{\inits{P.}\fnms{Peter}
    \snm{Sefton}\ead[label=ePS]{Peter.Sefton@uts.edu.au}\ead[label=oPS,orcid]{0000-0002-3545-944X}},
\author[MC]{\inits{M.}\fnms{Mercè}
    \snm{Crosas}\ead[label=eMC]{mcrosas@g.harvard.edu}\ead[label=oMC,orcid]{0000-0003-1304-1939}},
\author[LJC]{\inits{LJ.}\fnms{Leyla Jael}
    \snm{Castro}\ead[label=eLJC]{ljgarcia@zbmed.de}\ead[label=oLJC,orcid]{0000-0003-3986-0510}},
\author[FC]{\inits{F.}\fnms{Frederik}
    \snm{Coppens}\ead[label=eFC]{frcop@psb.vib-ugent.be}\ead[label=oFC,orcid]{0000-0001-6565-5145}},
\author[JMF]{\inits{JM.}\fnms{José M.}
    \snm{Fernández}\ead[label=eJMF]{jose.m.fernandez@bsc.es}\ead[label=oJMF,orcid]{0000-0002-4806-5140}},
\author[DG]{\inits{D.}\fnms{Daniel}
    \snm{Garijo}\ead[label=eDG]{dgarijo@fi.upm.es}\ead[label=oDG,orcid]{0000-0003-0454-7145}},
\author[BG]{\inits{B.}\fnms{Björn}
    \snm{Grüning}\ead[label=eBG]{gruening@informatik.uni-freiburg.de}\ead[label=oBG,orcid]{0000-0002-3079-6586}},
\author[MLR]{\inits{M.}\fnms{Marco}
    \snm{La Rosa}\ead[label=eMLR]{m@lr.id.au}\ead[label=oMLR,orcid]{0000-0001-5383-6993}},
\author[SL]{\inits{S.}\fnms{Simone}
    \snm{Leo}\ead[label=eSL]{simone.leo@crs4.it}\ead[label=oSL,orcid]{0000-0001-8271-5429}},
\author[EOC]{\inits{E.}\fnms{Eoghan}
    \snm{Ó Carragáin}\ead[label=eEOC]{eoghan.ocarragain@ucc.ie}\ead[label=oEOC,orcid]{0000-0001-8131-2150}},
\author[MP]{\inits{M.}\fnms{Marc}
    \snm{Portier}\ead[label=eMP]{marc.portier@vliz.be}\ead[label=oMP,orcid]{0000-0002-9648-6484}},
\author[AT]{\inits{A.}\fnms{Ana}
    \snm{Trisovic}\ead[label=eMP]{ana.trisovic@gmail.com}\ead[label=oAT,orcid]{0000-0003-1991-0533}},
\author[RO]{RO-Crate Community},
\author[PG]{\inits{P.}\fnms{Paul}
    \snm{Groth}\ead[label=ePG]{p.t.groth@uva.nl}\ead[label=oPG,orcid]{0000-0003-0183-6910}},
\author[CG]{\inits{C.}\fnms{Carole}
    \snm{Goble}\ead[label=eCG]{carole.goble@manchester.ac.uk}\ead[label=oCG,orcid]{0000-0003-1219-2137}}

\address[SSR1]{Department of Computer Science, 
    \orgname{The University of Manchester}, \cny{UK}\printead[presep={\\}]{eSSR,oSSR}}
\address[SSR2]{Informatics Institute, \orgname{University of Amsterdam},
    \cny{The Netherlands}}
\address[PS]{Faculty of Science, \orgname{University Technology Sydney},
    \cny{Australia}\printead[presep={\\}]{ePS,oPS}}
\address[MC]{Institute for Quantitative Social Science, 
    \orgname{Harvard University}, Cambridge, MA,  \cny{USA}\printead[presep={\\}]{oMC}}
\address[LJC]{\orgname{ZB MED Information Centre for Life Sciences}, 
    Cologne, \cny{Germany}\printead[presep={\\}]{oLJC}}
\address[FC]{\orgname{VIB-UGent Center for Plant Systems Biology}, 
    Gent, \cny{Belgium}\printead[presep={\\}]{oFC}}
\address[JMF]{\orgname{Barcelona Supercomputing Center}, 
    Barcelona, \cny{Spain}\printead[presep={\\}]{oJMF}}
\address[DG]{Ontology Engineering Group, 
    \orgname{Universidad Politécnica de Madrid}, Madrid, \cny{Spain}\printead[presep={\\}]{oDG}}
\address[BG]{Bioinformatics Group, Department of Computer Science,
    \orgname{Albert-Ludwigs-University Freiburg}, Freiburg, \cny{Germany}\printead[presep={\\}]{oBG}}
\address[MLR]{\orgname{PARADISEC}, Melbourne, \cny{Australia}\printead[presep={\\}]{oMLR}}
\address[SL]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA),
    \cny{Italy}\printead[presep={\\}]{oSL}}
\address[EOC]{\orgname{University College Cork},
    \cny{Ireland}\printead[presep={\\}]{oEOC}}
\address[MP]{\orgname{Vlaams Instituut voor de Zee}, Oostende,
    \cny{Belgium}\printead[presep={\\}]{oMP}}
\address[AT]{Institute for Quantitative Social Science,
    \orgname{Harvard University}, Cambridge, MA, \cny{USA}\printead[presep={\\}]{oAT}}
\address[RO]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
\address[PG]{Informatics Institute, \orgname{University of Amsterdam},
    \cny{The Netherlands}\printead[presep={\\}]{ePG,oPG}}
\address[CG]{Department of Computer Science, \orgname{The University of Manchester},
    \cny{UK}\printead[presep={\\}]{eCG,oCG}}

%\begin{review}{editor}
%\reviewer{\fnms{First} \snm{Editor}\ead[label=oe1,orcid]{}\printead[presep={\ }]{oe1}}
%\reviewer{\fnms{Second} \snm{Editor}\ead[label=oe2,orcid]{}\printead[presep={\ }]{oe2}}
%\end{review}
%\begin{review}{solicited}
%\reviewer{\fnms{First} \snm{Solicited reviewer}\ead[label=os1,orcid]{}\printead[presep={\ }]{os1}}
%\end{review}

\begin{abstract}
\markdownInput{../content/01.abstract.md}
\end{abstract}

\begin{keyword}
\kwd{Data publishing}
\kwd{Data packaging}
\kwd{FAIR}
\kwd{Linked Data}
\kwd{Metadata}
\kwd{Reproducibility}
\kwd{Research Object}
\end{keyword}

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Main text entry area:
\section{Introduction}

@@

\begin{figure}%[t!]
\includegraphics{content/images/ro-crate-overview.pdf}
\caption{Conceptual overview of RO-Crate. A \emph{Persistent
Identifier} (PID) \cite{doi:10.1371/journal.pbio.2001414} points to a
\emph{Research Object} (RO), which may be archived using different
packaging approaches like BagIt \cite{doi:10.17487/rfc8493}, OCFL \cite
{ocfl_2020}, git or ZIP. The RO is described within a \emph{RO-Crate
Metadata File}, providing identifiers for \emph{authors} using ORCID,
\emph{organisations} using Research Organization Registry (ROR) \cite
{doi:10.6087/kcse.192} and licences such as Creative Commons using SPDX
identifiers. The \emph{RO-Crate content} is further described with
additional metadata following a Linked Data approach. Data can be
embedded files and directories, as well as links to external Web
resources, PIDs and nested RO-Crates.}
\label{fig:conceptual}
\end{figure}

@@

\begin{figure}%[t!]
\includegraphics[width=0.9\textwidth]{content/images/ro-crate-uml.pdf}
\caption{Simplified UML class diagram of RO-Crate. The \emph
{RO-Crate Metadata File} conforms to a version of the specification;
and contains a JSON-LD graph \cite{sporny_2014} that describes the
entities that make up the RO-Crate. The \emph{RO-Crate Root Data
Entity} represent the Research Object as a dataset. The RO-Crate
aggregates \emph{data entities} (\texttt{hasPart}) which are further
described using \emph{contextual entities} (which may include
aggregated and non-aggregated data entities). Multiple types and
relations from Schema.org allow annotations to be more specific,
including figures, nested datasets, computational workflows, people,
organisations, instruments and places. Contextual entities not
otherwise cross-referenced from other entities' properties (\emph
{describes}) can be grouped under the root entity (\texttt{mentions}).}
\label{fig:uml}
\end{figure}

@@

%
\begin{figure}%[t]
\includegraphics{content/images/ro-crate-bco-sep-of-concerns.pdf}
\caption{Separation of Concerns in BCO RO-Crate. BioCompute
Object (IEEE2791) is a JSON file that structurally explains the purpose
and implementation of a computational workflow, for instance
implemented in Common Workflow Language (CWL), that installs the
workflow's software dependencies as Docker containers or BioConda
packages. An example execution of the workflow shows the different
kinds of result outputs, which may be external, using GitHub LFS \cite
{github-lfs} to support larger data. RO-Crate gathers all these local
and external resources, relating them and giving individual
descriptions, for instance permanent DOI identifiers for reused
datasets accessed from Zenodo, but also adding external identifiers to
attribute authors using ORCID or to identify which licences apply to
individual resources. The RO-Crate and its local files are captured in
a BagIt whose checksum ensures completeness, combined with Big Data Bag
\cite{doi:10.1109/BigData.2016.7840618} features to ``complete'' the
bag with large external files such as the workflow outputs.}
\label{fig:sep_concerns}
\end{figure}
%
\subsection{Digital humanities: Cultural heritage}% {#culturalheritage}

@@

\begin{figure}%[t!]
\includegraphics{content/images/data-commons-ro-crate-figure-5.pdf}
\caption{One aspect of Harvard Data Commons. Automatic
encapsulation and deposit of artefacts from data management tools used
during active research at the Harvard Dataverse repository.}
\label{fig:hdc}
\end{figure}

@@

%\markdownInput{../content/70.acknowledgements.md}

%%%% Appendix

%\newpage
%\appendix
%%\markdownInput{../content/80.formal-definition.md}
%\input{../content/80.formal-definition}

%\newpage
%\markdownInput{../content/89.community.md}

%%%%%%%%%%% The bibliography starts:
%\nocite{label}
%\bibliographystyle{ios1}
%\bibliography{ro-crate}

% Explicit bibliography from proof
%
% structpyb loaded by imikolaityte, 2021-12-28 10:21:28
%
\begin{thebibliography}{}

@@

\end{thebibliography}

\end{document}
